Black-hole ringdown as a probe of higher-curvature gravity theories

### 2022-04-05 papers/paper_alt_theor_bounds/paper_alt_theor_bounds.tex
hs: darboux refs and quartic eft text

Second, the scalar field does influence the response of BHs to linearized perturbations and thus
affects the BH's QNM spectra.
%
The coupling between scalar field and the Gauss-Bonnet invariant, which
manifests into a coupling between scalar perturbation and gravitational
perturbations of polar parity, which, e.g., breaks the equivalence
between the QNM spectra of polar and axial gravitational perturbations
(isospectrality~\cite{Chandrasekhar:1985kt}) of Schwarzschild BHs in GR.
%
The QNMs on nonrotating BHs in EdGB gravity were first computed in
Refs.~\cite{Pani:2009wy,Blazquez-Salcedo:2016enn} and these calculations were
extended to leading-order in BH spin in Ref.~\cite{Pierini:2021jxd}.

@@

We start by showing our results for EdGB gravity.
%
In Fig.~\ref{fig:edgb_exec_sum} we show the marginalized posteriors of the
coupling constant $\ell_{\rm GB}$, for GW150914 (top panel) and GW200129
(middle panel), with and without the spin corrections to the $(2,2,0)$ QNM.
%
The bottom panel panel shows the joint posterior combining both events.
%
We see that for both events events, the $N_{\rm max} = 0$ posteriors are
characterized by a peak away from zero.
%
This does not mean that we are inferring a deviation from GR from data.
%
Recall that the deviations from GR in the ParSpec framework are controlled
by $\gamma$, which here reads,
%
\begin{equation}
    \gamma_{\rm GB} = \left(\frac{c^2 \ell_{\rm GB}}{ G M_{\rm s}}\right)^4\,.
\end{equation}
%
As shown in Fig.~xxx, $\gamma_{\rm GB}$ does indeed have a posterior
distribution with largest support at zero, indicating consistency between data
and GR.
%
We also observe that the inclusion of the spin corrections (i.e., curves with $N_{\rm max} = 1$)
displaces the posteriors distributions towards smaller values of $\ell_{\rm GB}$.

@@

We now consider dCS gravity.
%
In Fig.~\ref{fig:dCS_exec_sum} we show the marginalized posterior distribution
function of $\ell_{\rm CS}$ for both events GW150914 (top panel) and GW200129
(middle panel), with different colors distinguishing cases with and without spin corrections to
dominant $(2,2,0)$~QNM.
%
Perhaps the most salient feature of these posteriors is that they are not
peaked at $\ell_{\rm CS}$, which may mislead us to believe that we are seeing a
deviation from GR. This is not the case and the situation is that same as the
one we discussed in the previous section for EdGB gravity.
%
We verified that
%
\begin{equation}
    \gamma_{\rm CS} = \left( \frac{c^2 \ell_{\rm CS}}{G M_{\rm s}} \right)^{4}\,,
\end{equation}
%
does indeed peak at zero indicating consistency with GR, similarly to what is shown in Fig.~\ref{} for $\gamma_{\rm GB}$.
%
We also see that in both cases, that the inclusion leading-order-in-spin
correction to the mode, displaces the posteriors towards smaller values of
$\ell_{\rm CS}$. This can be seen more evidently by looking at the location of
posterior peaks.
%
Finally, in the bottom panel, we show the combined result for both events.

@@

In Fig.~\ref{fig:cEFT_cdf} we show the CDF for
both events, calculated as already described for the EdGB and dCS cases.
%
We see that curves are very similar to those of dCS gravity for GW200129 (cf.
bottom panel in Fig.~\ref{fig:cEFT_cdf}).
%
Moreover, we find that for both events both the EFT~\eqref{eq:eft_bound} and
ParSpec~\eqref{eq:parspec_bound} bounds are satisfied for $\mathfrak{m} = M_f$
and $N_{\rm max} = 1/2$, allowing us to obtain the combined bound
%
\begin{equation}
    \ell_{\rm cEFT} \leqslant 51.3~\textrm{km},
\end{equation}
%
at 90\% credibility. As also happened for our study for dCS, the find that for
the cubic EFT that the ParSpec bound is violated by the $n = 1$ corrections to
the QNMs, implying that we cannot use the associated posterior distributions on
$\ell_{\rm cEFT}$ to draw any meaningful constraint on this parameter.

@@

@@

\begin{table}[h]
\begin{tabular}{l l c c c}
\hline
\hline
$N_{\rm max}$ & Event & EFT    & ParSpec & Constraint ($\gm = M_{f}$) \\
              &       & bound? & bound?  &                            \\
\hline
  & GW150914 & Yes & Yes  & $\ell_{\rm qEFT} \leqslant 51.7$~km \\
0 & GW200129 & Yes & Yes  & $\ell_{\rm qEFT} \leqslant 54.8$~km \\
  & Combined &     &      & \cellcolor{black!10}$\ell_{\rm qEFT} \leqslant 51.3$~km \\
\hline
  & GW150914 &  &     &    \\
1 & GW200129 &  &     &    \\
  & Combined &  &     &    \\
\hline
\hline
\end{tabular}
\caption{Detailed summary of our results the quartic EFT of GR for GW150914, GW200129, and
combined events using $\gm = M_{f}$, $\varepsilon = 1/2$ and quoting only 90\% credible results.
}
\label{tab:summary_qeft}
\end{table}

>>>>

Second, the scalar field does influence the response of BHs to linearized perturbations and thus
affects the BH's QNM spectra.
%
The coupling between scalar field and the Gauss-Bonnet invariant, which
result into a coupling between scalar perturbation and gravitational
perturbations of polar parity, which, e.g., breaks the equivalence
between the QNM spectra of polar and axial gravitational perturbations
(isospectrality~\cite{Chandrasekhar:1985kt}, see also Refs.~\cite{Glampedakis:2017rar,Lenzi:2021njy}) of Schwarzschild BHs in GR.
%
The QNMs on nonrotating BHs in EdGB gravity were first computed in
Refs.~\cite{Pani:2009wy,Blazquez-Salcedo:2016enn} and these calculations were
extended to leading-order in BH spin in Ref.~\cite{Pierini:2021jxd}.

@@

We start by showing our results for EdGB gravity.
%
In Fig.~\ref{fig:edgb_exec_sum} we show the marginalized posteriors of the
coupling constant $\ell_{\rm GB}$, for GW150914 (top panel) and GW200129
(middle panel), with and without the spin corrections to the $(2,2,0)$ QNM.
%
The bottom panel panel shows the joint posterior combining both events.
%
We see that for both events events, the $N_{\rm max} = 0$ posteriors are
characterized by a peak away from zero.
%
This does not mean that we are inferring a deviation from GR from data.
%
Recall that the deviations from GR in the ParSpec framework are controlled
by $\gamma$, which here reads,
%
\begin{equation}
    \gamma_{\rm GB} = \left(\frac{c^2 \ell_{\rm GB}}{ G M_{f}}\right)^4\,.
\end{equation}
%
As shown in Fig.~xxx, $\gamma_{\rm GB}$ does indeed have a posterior
distribution with largest support at zero, indicating consistency between data
and GR.
%
We also observe that the inclusion of the spin corrections (i.e., curves with $N_{\rm max} = 1$)
displaces the posteriors distributions towards smaller values of $\ell_{\rm GB}$.

@@

We now consider dCS gravity.
%
In Fig.~\ref{fig:dCS_exec_sum} we show the marginalized posterior distribution
function of $\ell_{\rm CS}$ for both events GW150914 (top panel) and GW200129
(middle panel), with different colors distinguishing cases with and without spin corrections to
dominant $(2,2,0)$~QNM.
%
Perhaps the most salient feature of these posteriors is that they are not
peaked at $\ell_{\rm CS}$, which may mislead us to believe that we are seeing a
deviation from GR. This is not the case and the situation is that same as the
one we discussed in the previous section for EdGB gravity.
%
We verified that
%
\begin{equation}
    \gamma_{\rm CS} = \left( \frac{c^2 \ell_{\rm CS}}{G M_{f}} \right)^{4}\,,
\end{equation}
%
does indeed peak at zero indicating consistency with GR, similarly to what is shown in Fig.~\ref{} for $\gamma_{\rm GB}$.
%
We also see that in both cases that the inclusion leading-order-in-spin
correction to the mode displaces the posteriors towards smaller values of
$\ell_{\rm CS}$. This can be seen more evidently by looking at the location of
posterior peaks.
%
Finally, in the bottom panel, we show the combined result for both events.

@@

In Fig.~\ref{fig:cEFT_cdf} we show the CDF for
both events, calculated as already described for the EdGB and dCS cases.
%
We see that curves are very similar to those of dCS gravity for GW200129 (cf.
bottom panel in Fig.~\ref{fig:cEFT_cdf}).
%
Moreover, we find that for both events both the EFT~\eqref{eq:eft_bound} and
ParSpec~\eqref{eq:parspec_bound} bounds are satisfied for $\mathfrak{m} = M_f$
and $N_{\rm max} = 1/2$, allowing us to obtain the combined bound
%
\begin{equation}
    \ell_{\rm cEFT} \leqslant 51.3~\textrm{km},
\end{equation}
%
at 90\% credibility. As also happened for our study for dCS, the find that for
the cubic EFT that the ParSpec bound is violated by the $n = 1$ corrections to
the QNMs, implying that we cannot use the associated posterior distributions on
$\ell_{\rm cEFT}$ to draw any meaningful constraint on this parameter.
%
We summarize our results in Table~\ref{tab:summary_ceft}.

@@

In this theory, we considered only $N_{\rm max} = 0$. We found that the
inclusion of the spin corrections yields waveforms which for $\ell_{\rm qEFT} \gtrsim 50$~km
with ringdowns portions of the larger than the inspiral-plunge part \dots
%
To avoid this issue we lowered the value of $\ell^{\rm max}_{\rm qEFT}$, but at doing so we obtained which
were essentially uniform. Thus, we do not quote any results for $N_{\rm max} = 1$.
%
Table~\ref{tab:summary_qeft} summarizes our findings for the quartic EFT of GR.

@@

\begin{table}[h]
\begin{tabular}{l l c c c}
\hline
\hline
$N_{\rm max}$ & Event & EFT    & ParSpec & Constraint ($\gm = M_{f}$) \\
              &       & bound? & bound?  &                            \\
\hline
  & GW150914 & Yes & Yes  & $\ell_{\rm qEFT} \leqslant 51.7$~km \\
0 & GW200129 & Yes & Yes  & $\ell_{\rm qEFT} \leqslant 54.8$~km \\
  & Combined &     &      & \cellcolor{black!10}$\ell_{\rm qEFT} \leqslant 51.3$~km \\
% \hline
%   & GW150914 &  &     &    \\
% 1 & GW200129 &  &     &    \\
%   & Combined &  &     &    \\
\hline
\hline
\end{tabular}
\caption{Detailed summary of our results the quartic EFT of GR for GW150914, GW200129, and
combined events using $\gm = M_{f}$, $\varepsilon = 1/2$ and quoting only 90\% credible results.
}
\label{tab:summary_qeft}
\end{table}
